---
title: "Assignment-2"
author: "Kazim shah"
date: "9/11/2020"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)
library(lubridate)
library(skimr)
library(broom)
library(mgcv)
library(visreg)
library(zoo)
library(data.table)
library(weathermetrics)
```

#Importing weather and hospital data and storing them in working folder for further manipulation
```{r}
#We can download hospital data from below link
hospital_data_raw <- read.csv(file = 'http://bit.ly/2nkCUEh')
#write.csv(hospital_data_raw,'Hospital_data.csv')

hospital_data<-read_csv('govhack3.csv')

#Weather data is pulled from NOAA website, file 2277258.csv is also submitted with assignment
weather_st_data_raw<-read.csv('2277258.csv')

```
#Checking number of rows in the weather data and time period

```{r}

print(paste("Total number of rows: ",nrow(weather_st_data_raw)))


weather_st_data_date<-transform(weather_st_data_raw,DATE=as.Date (DATE))
date_data<-arrange(weather_st_data_date,(DATE))

print(paste("The first entry recorded on",first(date_data$DATE)))
print(paste("The last entry recorded on",last(date_data$DATE)))

```

#Making hospital data tidy
#Code referenced from assignment 1 as same data set is being used  

```{r}
#Seperating first row of the data to use further in the code

hospital_list<-'govhack3.csv'
#hospital_list<-'Hospital_data.csv'
row_1 <- read_csv(hospital_list, col_names=FALSE, n_max = 1)
list_of_hospitals<-c(row_1)
list_of_hospitals<-list_of_hospitals[!is.na(list_of_hospitals)]

#Making 2nd row of the dataset the header
#hospital_data_v2<-read.csv('Hospital_data.csv',skip=1)

hospital_data_v2<-read.csv('govhack3.csv',skip=1)
write.csv(hospital_data_v2,'data_modified.csv')
data_new<-read_csv('data_modified.csv')

#Formating hospital names so that it can be combined in the same format with rest of the data 
new_hospital_df<-data.frame(list_of_hospitals)
hospitals<-t(new_hospital_df)
h_1<-rep(hospitals[1,],365)
h_2<-rep(hospitals[2,],365)
h_3<-rep(hospitals[3,],365)
h_4<-rep(hospitals[4,],365)
h_5<-rep(hospitals[5,],365)
h_6<-rep(hospitals[6,],365)
h_7<-rep(hospitals[7,],365)
h_8<-rep(hospitals[8,],365)
h_9<-rep(hospitals[9,],365)

h_combo<-data.frame(h_1,h_2,h_3,h_4,h_5,h_6,h_7,h_8,h_9)
df2_hospitlas<-data.frame(Hospital=unlist(h_combo,use.names=FALSE))

data_tidy_admission <- hospital_data_v2 %>% 
  gather(key = "Admissions_new", 
         value = "Admissions", 
         starts_with("Admission"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_attendence <- hospital_data_v2 %>% 
  gather(key = "Attendance_new", 
         value = "Attendance", 
         starts_with("Attendance"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_1 <- hospital_data_v2 %>% 
  gather(key = "Triage_1", 
         value = "Tri_1_cases", 
         starts_with("Tri_1"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_2 <- hospital_data_v2 %>% 
  gather(key = "Triage_2", 
         value = "Tri_2_cases", 
         starts_with("Tri_2"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_3 <- hospital_data_v2 %>% 
  gather(key = "Triage_3", 
         value = "Tri_3_cases", 
         starts_with("Tri_3"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_4 <- hospital_data_v2 %>% 
  gather(key = "Triage_4", 
         value = "Tri_4_cases", 
         starts_with("Tri_4"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_5 <- hospital_data_v2 %>% 
  gather(key = "Triage_5", 
         value = "Tri_5_cases", 
         starts_with("Tri_5"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_modi_date<-data_tidy_admission["Date"]
data_modi_atte<-data_tidy_attendence[c("Attendance")]
data_modi_admi<-data_tidy_admission[c("Admissions")]
data_modi_tria_1<-data_tidy_triage_1[c("Tri_1_cases")]
data_modi_tria_2<-data_tidy_triage_2[c("Tri_2_cases")]
data_modi_tria_3<-data_tidy_triage_3[c("Tri_3_cases")]
data_modi_tria_4<-data_tidy_triage_4[c("Tri_4_cases")]
data_modi_tria_5<-data_tidy_triage_5[c("Tri_5_cases")]
Hospital_data_tidy=data.frame(data_modi_date,df2_hospitlas,data_modi_atte,data_modi_admi,data_modi_tria_1,data_modi_tria_2,data_modi_tria_3,data_modi_tria_4,data_modi_tria_5)
#Rearranging Date column for final view

Hospital_data_tidy<-transform(Hospital_data_tidy, Hospital=as.character(Hospital),Admissions=as.numeric(Admissions),Attendance=as.numeric(Attendance),Tri_1_cases=as.numeric(Tri_1_cases),Tri_2_cases=as.numeric(Tri_2_cases),Tri_3_cases=as.numeric(Tri_3_cases),Tri_5_cases=as.numeric(Tri_5_cases))

Hospital_data_tidy[is.na(Hospital_data_tidy)] <- 0

#Selecting Royal perth hospital as it closest to weather station
Hospital_select<-Hospital_data_tidy %>% filter(Hospital=="Royal Perth Hospital")

Hospital_select<-Hospital_select %>%
  mutate(Date=seq(from = as.Date("2013-07-01"), to = as.Date("2014-06-30"), by = 'day'))

Hospital_select

```

#Creating linear model for ED demands with date

```{r}
#Taking ED demand average of all the hospitals in perth area
skim(Hospital_select)

Hospital_select %>%
  ggplot(aes(x=Date,y=Attendance))+geom_point()+geom_smooth(method = lm)+ggtitle("Royal perth ED Demand Data")
  
```

```{r}
#Adding week day number which will be used further
Hospital_select_week<-Hospital_select%>% mutate(weekday=wday(Hospital_select$Date))
Hospital_select_lm<-transform(Hospital_select_week,Date=as.numeric(Date))
#Linear regression model with date
lmfit<-lm(Attendance~Date,data =Hospital_select_lm)
lmfit%>%summary
```

```{r}
lmfit %>%
  tidy
```

```{r}
results <- lmfit %>% 
  augment
```

```{r}
lmfit %>% 
  glance
```

```{r}
results %>% 
  ggplot(aes(sample = .std.resid)) +
  geom_qq() + 
  geom_qq_line(col = 'steelblue')+ggtitle('QQ plot for Standard residuals')
```

```{r}
results %>% 
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_quantile() +
  geom_smooth(colour = 'firebrick')+ggtitle('Fitted vs Residuals')
```

```{r}
results %>% 
  ggplot(aes(x = Attendance, y = .resid)) +
  geom_point() + 
  geom_quantile() +
  geom_smooth(colour = 'firebrick') 
```

```{r}
results %>% 
  ggplot(aes(x = .hat, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0,
             linetype = 'dotted')
```

```{r}
results %>% 
  mutate(big_hat = .hat > 0.03) %>% 
  ggplot(aes(x = .hat, y = .cooksd)) +
  geom_point(aes(color = big_hat)) + 
  geom_hline(yintercept = 0,
             linetype = 'dotted')
```


#Fitting GAM model
```{r}
Hospital_gam <- gam(Attendance ~ s(Date),
           family=Gamma(link=log),
           data=Hospital_select_lm)
summary(Hospital_gam)
```

```{r}
plot(Hospital_gam)
```

```{r}

Hospital_select_lm %>%
  ggplot(aes(x=Date,y=Attendance))+geom_point()+geom_smooth(method = 'gam',formula=y~s(x))+ggtitle("Royal perth ED Demand Data using GAM")


#ggplot(Sample_data, aes(x, y)) + geom_point() + geom_smooth(method = "gam", formula = y ~s(x))
```


```{r}
plot(Hospital_gam, residuals = TRUE)
```

```{r}
#Incorporating weely seasonality 
#Hospital_select_lm<-transform(Hospital_select,Date=as.Date(Date))

#GAM with date and weekday
Hospital_gam_1 <- gam(Attendance ~ s(Date)+s(weekday,k=2),
           family=Gamma(link=log),
           data=Hospital_select_lm)
summary(Hospital_gam_1)


AIC(lmfit,Hospital_gam,Hospital_gam_1)

```

```{r}
plot(Hospital_gam_1)
```

```{r}
plot(Hospital_gam_1, residuals = TRUE)
```



#Task4: Calculating EHF


#Making weather station data tidy
```{r}
#Removing un-necessary columns from weather data set
#weather_st_data_raw
#Filtering perth weather station
weather_st_perth<-weather_st_data_raw %>% filter(STATION=="ASN00009021")
weather_st_perth <- weather_st_perth[-c(3:5,7:20,22,24,26,28)]
weather_st_perth<-transform(weather_st_perth,STATION=as.character(STATION),NAME=as.character(NAME), Date=as.Date (DATE),PRCP=as.double(PRCP),TAVG=as.double(TAVG),TMAX=as.double(TMAX),TMIN=as.double(TMIN))
weather_st_perth <- weather_st_perth[-c(3)]
#weather_st_data_tidy<-weather_st_data_raw %>%
 # group_by(DATE) %>%
  #summarise(Avg_Temperature_F=mean(TAVG,na.rm=TRUE),Avg_PRCP_mm=mean(PRCP,na.rm=TRUE),Max_Temp_F=mean(TMAX,na.rm=TRUE),Min_temp_F=mean(TMIN,na.rm=TRUE))
#weather_st_data_tidy
weather_st_perth
```


```{r}
#Creating 3-day rolling average to create EHF

weather_st_perth <- weather_st_perth %>% mutate(EHF= rollmeanr(weather_st_perth[,4],3,fill=0))
weather_st_perth %>%
  ggplot(aes(x=Date,y=EHF))+geom_point()+geom_smooth(method = lm)+ggtitle('EHF yearly plot')

```


```{r}
#Merging EHF values with ED demand data and ploting data
weather_st_perth_EHF<-weather_st_perth[,c(7,8)]

setDT(weather_st_perth_EHF)
setDT(Hospital_select)

merge_data<-Hospital_select[weather_st_perth_EHF, on = c('Date')]
```


```{r}
#Creating linear model with EHF values
merge_data_lm<-merge_data%>% mutate(weekday=wday(merge_data$Date))
merge_data_lm<-transform(merge_data_lm,Date=as.numeric(Date))

merge_data_lm$EHF_c <- fahrenheit.to.celsius(merge_data_lm$EHF)
#Replacing 0 with mean value
merge_data_lm$EHF_c[1]<-mean(merge_data_lm$EHF_c)
merge_data_lm$EHF_c[2]<-mean(merge_data_lm$EHF_c)
merge_data_lm %>%
  ggplot(aes(x=EHF_c,y=Attendance))+geom_point()+geom_smooth(method = lm)+ggtitle('Attendence Vs EHF_celsius')


lmfit_3<-lm(Attendance~Date+weekday+EHF_c,data =merge_data_lm)
lmfit_3%>%summary



```

```{r}
#Linear model with EHF_c
lmfit_4<-lm(Attendance~EHF_c,data =merge_data_lm)
lmfit_4%>%summary

```



```{r}
merge_data_lm %>%
  ggplot(aes(x=EHF_c,y=Attendance))+geom_point()+geom_smooth(method = 'gam',formula=y~s(x))+ggtitle("ED Demands Vs EHF")
```


```{r}
Hospital_gam_ehf <- gam(Attendance ~ s(EHF_c),
           family=Gamma(link=log),
           data=merge_data_lm)
summary(Hospital_gam_ehf)


```

```{r}
merge_data_lm %>%
  ggplot(aes(x=Date,y=EHF_c))+geom_point()+geom_smooth(method = 'gam',formula=y~s(x))+ggtitle("GAM with EHF")
```

```{r}
plot(Hospital_gam_ehf, residuals = TRUE)
```

```{r}
Hospital_gam_ehf_1 <- gam(Attendance ~ s(Date)+s(weekday,k=2)+s(EHF_c,k=2),
           family=Gamma(link=log),
           data=merge_data_lm)
summary(Hospital_gam_ehf_1)


AIC(lmfit_3,lmfit_4, Hospital_gam_ehf,Hospital_gam_ehf_1)
```
#Extra weather features

```{r}
Hospital_select_weather<-Hospital_select[,c(1,3)]

setDT(weather_st_perth)
setDT(Hospital_select_weather)

merge_data_1<-weather_st_perth[Hospital_select_weather, on = c('Date')]

#merge_data_lm$EHF_c <- fahrenheit.to.celsius(merge_data_lm$EHF)

merge_data_1$EHF_c <- fahrenheit.to.celsius(merge_data_1$EHF)
merge_data_1$TVG_c <- fahrenheit.to.celsius(merge_data_1$TAVG)
merge_data_1$TMAX_c <- fahrenheit.to.celsius(merge_data_1$TMAX)
merge_data_1$TMIN_c <- fahrenheit.to.celsius(merge_data_1$TMIN)
#Replace 0 values with mean values
merge_data_1$EHF_c[1]<-mean(merge_data_1$EHF_c)
merge_data_1$EHF_c[2]<-mean(merge_data_1$EHF_c)
```


```{r}
merge_data_1 %>%
  ggplot(aes(x=TVG_c,y=Attendance))+geom_point()+geom_smooth(method = lm)+ggtitle("Royal perth ED Demand Data")
merge_data_1 %>%
  ggplot(aes(x=PRCP,y=Attendance))+geom_point()+geom_smooth(method = lm)+ggtitle("Royal perth ED Demand Data")


lmfit_5<-lm(Attendance~EHF+PRCP,data =merge_data_1)
summary(lmfit_5)

```


```{r}
lmfit_6<-lm(Attendance~EHF+PRCP+TVG_c+TMAX_c+TMIN_c,data =merge_data_1)
summary(lmfit_6)


```










