---
title: "Assignment_1"
author: "Kazim shah"
date: "8/7/2020"
output: html_document
---
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)
library(fitdistrplus)
library(magrittr)
```

######kindly run all markdown cells.##########

Section 1
Q 1.1:Reading file as dataframe and displaying first few lines of the data set
```{r}
data <- read.csv(file = 'http://bit.ly/2nkCUEh')
head(data)
```

Q 1.2.1: How many rows and columns in the data
```{r}
print(paste("Total number of rows: ",nrow(data)))
print(paste("Total number of columns: ", ncol(data)))
```

Q 1.2:Counting number of hospitals in the dataset from the first row of the data
```{r}
header<-colnames(data)
sum_header<-str_count(header,"X")
print(paste("The total number of Hospitals in the data set are",sum(sum_header==0)))
```

Q 1.3: Checking data types of the data

kindly run all modlues

```{r}
#Seperating first row of the data to use further in the code
data_test<-'govhack3.csv'
row_1 <- read_csv(data_test, col_names=FALSE, n_max = 1)
list_of_hospitals<-c(row_1)
list_of_hospitals<-list_of_hospitals[!is.na(list_of_hospitals)]
```


```{r}
#Making 2nd row of the dataset the header
row_2<-read_csv('govhack3.csv',skip=1)
write.csv(row_2,'data_modified.csv')
data_new<-read_csv('data_modified.csv')
head(data_new)
```


```{r}
#printing data types
str(data_new)
```

Q 1.4: Checking Duration of time over which data is collected
```{r}
print(paste("The first entry recorded on",first(data_new$Date)))
print(paste("The last entry recorded on",last(data_new$Date)))
print(paste("So the total duration of the dataset is 1 year"))
```

Section 2

Q2.1: Printing list of hospitals after cleaning up data

```{r}
new_hospital_df<-data.frame(list_of_hospitals)
hospitals<-t(new_hospital_df)
hospitals
```

```{r}
conbo<-data.frame(data_new[2],h_1,h_2)
h_1<-rep(hospitals[1,],365)
h_2<-rep(hospitals[2,],365)
h_3<-rep(hospitals[3,],365)
h_4<-rep(hospitals[4,],365)
h_5<-rep(hospitals[5,],365)
h_6<-rep(hospitals[6,],365)
h_7<-rep(hospitals[7,],365)
h_8<-rep(hospitals[8,],365)
h_9<-rep(hospitals[9,],365)

h_combo<-data.frame(h_1,h_2,h_3,h_4,h_5,h_6,h_7,h_8,h_9)
df2_hospitlas<-data.frame(Hospital=unlist(h_combo,use.names=FALSE))
```



Q 2.2 (a): Does each variable has its own column?
No, as the data is gathered row wise with respect to date, variable names are same for each hospital so each variable is repeating itself for each hospital.
(b) Does each observation has its own row?
No, we can see that one observation is for one hospital, so observations for each hospital are in the same row.  
(c) Does each value has its own cell?
Yes.

Q 2.3: Use Spreadin/gathering to make data tidy
```{r}
data_tidy_admission <- data_new %>% 
  gather(key = "Admissions_new", 
         value = "Admissions", 
         starts_with("Admission"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_attendence <- data_new %>% 
  gather(key = "Attendance_new", 
         value = "Attendance", 
         starts_with("Attendance"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_1 <- data_new %>% 
  gather(key = "Triage_1", 
         value = "Tri_1_cases", 
         starts_with("Tri_1"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_2 <- data_new %>% 
  gather(key = "Triage_2", 
         value = "Tri_2_cases", 
         starts_with("Tri_2"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_3 <- data_new %>% 
  gather(key = "Triage_3", 
         value = "Tri_3_cases", 
         starts_with("Tri_3"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_4 <- data_new %>% 
  gather(key = "Triage_4", 
         value = "Tri_4_cases", 
         starts_with("Tri_4"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_tidy_triage_5 <- data_new %>% 
  gather(key = "Triage_5", 
         value = "Tri_5_cases", 
         starts_with("Tri_5"), 
         na.rm = FALSE) %>% 
  replace_na(list(cases = 0))

data_modi_date<-data_tidy_admission["Date"]
data_modi_atte<-data_tidy_attendence[c("Attendance")]
data_modi_admi<-data_tidy_admission[c("Admissions")]
data_modi_tria_1<-data_tidy_triage_1[c("Tri_1_cases")]
data_modi_tria_2<-data_tidy_triage_2[c("Tri_2_cases")]
data_modi_tria_3<-data_tidy_triage_3[c("Tri_3_cases")]
data_modi_tria_4<-data_tidy_triage_4[c("Tri_4_cases")]
data_modi_tria_5<-data_tidy_triage_5[c("Tri_5_cases")]
data_new_tidy=data.frame(data_modi_date,df2_hospitlas,data_modi_atte,data_modi_admi,data_modi_tria_1,data_modi_tria_2,data_modi_tria_3,data_modi_tria_4,data_modi_tria_5)
data_new_tidy
```

Q2-3: Variables Data types:

```{r}
data_new_tidy<-transform(data_new_tidy, Hospital=as.character(Hospital),Admissions=as.numeric(Admissions),Attendance=as.numeric(Attendance),Tri_1_cases=as.numeric(Tri_1_cases),Tri_2_cases=as.numeric(Tri_2_cases),Tri_3_cases=as.numeric(Tri_3_cases),Tri_5_cases=as.numeric(Tri_5_cases))

#as.Date(data_new_tidy[1],"08/16/1975")

#data_new_tidy[1]<-strptime(data_new_tidy[1],"%B, %d, %Y")

str(data_new_tidy)

```

Q2-4: Remove N/A Values:

```{r}
data_new_tidy[is.na(data_new_tidy)] <- 0
```

Q-3.1: Select a Hospital:
```{r}
Hospital_select<-data_new_tidy %>% filter(Hospital=="Royal Perth Hospital") 
print(paste("Hospital Name: ",unique(Hospital_select$Hospital)))
print(paste("Total Number of People Attended ED in 1 year: ",sum(Hospital_select$Admissions)))
print(paste("Total Number of people Admitted in ED in i year",sum(Hospital_select$Attendance)))
```

Q-3.2: Comparison of ED demands across the year

```{r}
Hospital_select$Week <- 1:nrow(Hospital_select) 
Hospital_select$Week <- ceiling((Hospital_select$Week/7))
#values = seq(from = as.Date("2013-07-01"), to = as.Date("2014-06-30"), by = 'day')
#Hospital_select<-merge(Hospital_select, values, all = T)

#Hospital_select <- Hospital_select[, c(10, 2, 3, 4, 5,6,7,8,9)]
#Hospital_select<-Hospital_select %>% 
 # rename(
  #  Date = y
   #   )

#plot(Hospital_select$Week,Hospital_select$Attendance)
#ggplot(Hospital_select, aes(Date, Attendance))+ geom_jitter()

ggplot(Hospital_select, aes(x = Week, y = Attendance)) +geom_line() + geom_point()  +ggtitle("Average Hospital Attendence per week")+geom_vline(xintercept = 18, linetype="dotted",color = "blue", size=1.5)+geom_vline(xintercept = 39, linetype="dotted",color = "red", size=1.5) 



#ggplot(Hospital_select, aes(x=Date, y=Attendance)) + geom_bar(stat="identity")



```

Q-3.3: How do the ED demands change during a week? Show it visually

```{r}

ggplot(Hospital_select, aes(x=Week, y=Attendance,group=Week)) + 
  geom_boxplot()+ggtitle("Average Hospital Attendence per week")

```

Q-3.4: Possion Distribution of randomly 200 samples:

```{r}

#Running Cullen and Frey graph for distribution selection
Hospital_select %$% 
  descdist(Attendance, boot = 100)

Hospital_select[sample(nrow(Hospital_select), 200),]


#Pois_dis <- Hospital_select$Attendance 
#fitp <- fitdist(Pois_dis,"pois")
#summary(fitp)
#plot(fitp)
#fitnb <- fitdist(number,"nbinom")
#summary(fitnb)
#plot(fitnb)

```


Q 4-1: Fitting Distributions:
```{r}
Pois_dis_1 <- Hospital_select$Tri_2_cases 

estimate <- Hospital_select %$%
  fitdist(data = Tri_2_cases, 
          distr = "pois")
estimate %>% 
  plot

fitp_1 <- fitdist(Pois_dis_1,"pois")
summary(fitp_1)
plot(fitp_1)
fitp_1 %>% 
  plot
fitnb <- fitdist(Pois_dis_1,"nbinom")
summary(fitnb)
plot(fitnb)
fitnb %>% 
  plot
```

Q 4-2 (a): Log-liklihood of two fitted distributions:
```{r}
Pois_ll<-logLik(fitp_1)
nb_ll<-logLik(fitnb)
ratio=Pois_ll/nb_ll
ratio
```

Q-4-2(b): Which distribution fit the data better:
Log likelihood ratio comes out to be 1.01425. This means that poison distribution model is 1.014 times more suitable against negative binomial distribution for projecting the data

Q-5: Fitting Distributions:
Mentions in report

Q-6: Ethics:
Mentions in report

Q-7: Reflection:
Mentions in report








